Decomposition-based Job-shop Scheduling with Constrained Clustering

### 2021-11-21 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
minor text modifications suggested by Reviewer 2 applied

Let us illustrate the problem on the small JSP instance specified in Table~\ref{tab1}, which provides parameters for $3$ jobs and $3$ machines. 
The rows list operations along with their respective machines and processing times; e.g., the third operation of the first job, $O_{1,3}$, takes $5$ time units for execution by machine~$1$.
The minimum makespan happens to be $20$, which matches the sum $9+6+5$ of processing times for the operations $O_{1,1}$, $O_{2,2}$, and $O_{3,3}$ executed by machine~$2$.
While there are plenty, i.e., $234$, optimal schedules with makespan $20$,
they agree on the execution orders $(O_{3,1}, O_{1,3}, O_{2,3})$ and $(O_{1,1}, O_{2,2}, O_{3,3})$
of operations processed by machine~$1$ or~$2$, respectively.
The two feasible execution orders for machine $3$ are
$(O_{2,1}, O_{3,2}, O_{1,2})$ and
$(O_{2,1}, O_{1,2}, O_{3,2})$.
In both cases, the earliest eligible starting times for the operations
$O_{1,1}$, $O_{1,2}$, and $O_{1,3}$ of job $J_1$ are
$0$, $9$, and $12$,
as well as
$0$, $9$, and $17$ for $O_{2,1}$, $O_{2,2}$, and $O_{2,3}$
belonging to the job~$J_2$.
Regarding the operations of~$J_3$,
the earliest starting time for $O_{3,1}$ is~$0$,
$O_{3,3}$ can only be started at time~$15$ because its machine~$1$
is occupied before,
while either $4$ or $12$ is the earliest eligible starting time for $O_{3,2}$,
depending on whether it is processed directly after completing $O_{3,1}$
(and $O_{2,1}$ on its machine~$3$) or waits for the completion of $O_{1,2}$
on machine~$3$.
The latter option lets machine~$3$ idle unnecessarily 
and may seem less attractive,
yet it does not stretch the resulting makespan beyond time~$20$. 
%
% In this table, rows correspond to operations, whereas columns refer to the name of the operation, the machine assigned to an operation, and the processing time of a particular operation. For instance, in the third row of the table, $O_{1,3}$ is the third operation of the first job that will be executed by machine $1$ with processing time $5$. 
%
\begin{table}[t]
%  \centering
\begin{minipage}{.56\textwidth}
%\fontsize{12}{9}
\setlength{\tabcolsep}{5.0pt}
\centering
 % \begin{center}
    \caption{A sample JSP instance.}
    \label{tab1}
 %\begin{adjustwidth}{0.0cm}{}
    \begin{tabular}{c c c}
      \textbf{Operation} & \textbf{Machine} & \textbf{Processing Time} \\
      \hline
								\\
      $O_{1,1}$  & $2$  &  $9$  	\\
      $O_{1,2}$  & $3$  &  $3$  	\\
      $O_{1,3}$  & $1$  &  $5$ 	\\
								\\
      $O_{2,1}$  & $3$  & $4$  	\\
      $O_{2,2}$  & $2$  & $6$  	\\
      $O_{2,3}$  & $1$  & $2$  	\\
								\\
      $O_{3,1}$  & $1$  & $4$  	\\
      $O_{3,2}$  & $3$  & $3$  	\\
      $O_{3,3}$  & $2$  & $5$  	\\
    \end{tabular}
%\end{adjustwidth}
%  \end{center}
\end{minipage}
\begin{minipage}{.4\textwidth}
  \setlength{\tabcolsep}{5.0pt}
  \centering
 % \begin{center}
    \caption{Possible decomposition.}
    \label{tab2}
 %\begin{adjustwidth}{0.0cm}{}
    \begin{tabular}{c  c }
      \textbf{Operation} & \textbf{Time Window}  \\
      \hline
							\\
      $O_{1,1}$  & $1$    	\\
      $O_{1,2}$  & $1$    	\\
      $O_{2,1}$  & $1$    	\\
      $O_{2,2}$  & $1$   	\\
      $O_{3,1}$  & $1$		\\
					    		\\
      $O_{1,3}$  & $2$  	\\
      $O_{2,3}$  & $2$		\\
      $O_{3,2}$  & $2$    	\\
      $O_{3,3}$  & $2$    	\\
      \\
    \end{tabular}
%\end{adjustwidth}
  %\end{center}
\end{minipage}
\end{table}

@@

The clustering algorithms are unsupervised learning methods which goal is to partition a set of objects into (disjoint) subsets, called clusters, such that each cluster contains the most similar objects. 
The similarity of objects is determined by some measure, like Euclidian distance, using features of each object, such as features of operations extracted in the previous section. 
However, direct application of clustering algorithms, like K-Means \cite{Forgy1965ClusterAO}, in scheduling is limited since the partitioning cannot preserve the sequences of operations in a job. As a result, a clustering algorithm might place $O_{1,1}$ and $O_{1,3}$ in one cluster and $O_{1,2}$ in another. As a result, the sequence of sub-problems becomes inconsistent and no schedule can be found.

@@

\section{Related Work}
\label{sec:literature}
In this section, we review some of the most related work using data mining approaches employed to solve scheduling problems. Over the last decades, solving the scheduling problem has attracted the interest of many researchers.The main idea of data mining in these applications is to explore the patterns from the obtained solutions by some heuristics and then develop decision rules that approximate the order of the operations in each machine \cite{ismail2012production}. More precisely, it aims to get an order of all operations in the given instance according to their characteristics.

>>>>

Let us illustrate the problem on the small JSP instance specified in Table~\ref{tab1}, which provides parameters for $3$ jobs and $3$ machines. 
The rows list operations along with their respective machines and processing times, e.g., the third operation of the first job, $O_{1,3}$, takes $5$ time units for execution by machine~$1$.
The minimum makespan happens to be $20$, which matches the sum $9+6+5$ of processing times for the operations $O_{1,1}$, $O_{2,2}$, and $O_{3,3}$ executed by machine~$2$.
While there are plenty, i.e., $234$, optimal schedules with makespan $20$,
they agree on the execution orders $(O_{3,1}, O_{1,3}, O_{2,3})$ and $(O_{1,1}, O_{2,2}, O_{3,3})$
of operations processed by machine~$1$ or~$2$, respectively.
The two feasible execution orders for machine $3$ are
$(O_{2,1}, O_{3,2}, O_{1,2})$ and
$(O_{2,1}, O_{1,2}, O_{3,2})$.
In both cases, the earliest eligible starting times for the operations
$O_{1,1}$, $O_{1,2}$, and $O_{1,3}$ of job $J_1$ are
$0$, $9$, and $12$,
as well as
$0$, $9$, and $17$ for $O_{2,1}$, $O_{2,2}$, and $O_{2,3}$
belonging to the job~$J_2$.
Regarding the operations of~$J_3$,
the earliest starting time for $O_{3,1}$ is~$0$,
$O_{3,3}$ can only be started at time~$15$ because its machine~$1$
is occupied before,
while either $4$ or $12$ is the earliest eligible starting time for $O_{3,2}$,
depending on whether it is processed directly after completing $O_{3,1}$
(and $O_{2,1}$ on its machine~$3$) or waits for the completion of $O_{1,2}$
on machine~$3$.
The latter option lets machine~$3$ idle unnecessarily 
and may seem less attractive,
yet it does not stretch the resulting makespan beyond time~$20$. 
%
% In this table, rows correspond to operations, whereas columns refer to the name of the operation, the machine assigned to an operation, and the processing time of a particular operation. For instance, in the third row of the table, $O_{1,3}$ is the third operation of the first job that will be executed by machine $1$ with processing time $5$. 
%
\begin{table}[t]
%  \centering
\begin{minipage}{.56\textwidth}
%\fontsize{12}{9}
\setlength{\tabcolsep}{5.0pt}
\centering
 % \begin{center}
    \caption{A sample JSP instance.}
    \label{tab1}
 %\begin{adjustwidth}{0.0cm}{}
    \begin{tabular}{c c c}
      \textbf{Operation} & \textbf{Machine} & \textbf{Processing Time} \\
      \hline
								\\
      $O_{1,1}$  & $2$  &  $9$  	\\
      $O_{1,2}$  & $3$  &  $3$  	\\
      $O_{1,3}$  & $1$  &  $5$ 	\\
								\\
      $O_{2,1}$  & $3$  & $4$  	\\
      $O_{2,2}$  & $2$  & $6$  	\\
      $O_{2,3}$  & $1$  & $2$  	\\
								\\
      $O_{3,1}$  & $1$  & $4$  	\\
      $O_{3,2}$  & $3$  & $3$  	\\
      $O_{3,3}$  & $2$  & $5$  	\\
    \end{tabular}
%\end{adjustwidth}
%  \end{center}
\end{minipage}
\begin{minipage}{.4\textwidth}
  \setlength{\tabcolsep}{5.0pt}
  \centering
 % \begin{center}
    \caption{Possible decomposition.}
    \label{tab2}
 %\begin{adjustwidth}{0.0cm}{}
    \begin{tabular}{c  c }
      \textbf{Operation} & \textbf{Time Window}  \\
      \hline
							\\
      $O_{1,1}$  & $1$    	\\
      $O_{1,2}$  & $1$    	\\
      $O_{2,1}$  & $1$    	\\
      $O_{2,2}$  & $1$   	\\
      $O_{3,1}$  & $1$		\\
					    		\\
      $O_{1,3}$  & $2$  	\\
      $O_{2,3}$  & $2$		\\
      $O_{3,2}$  & $2$    	\\
      $O_{3,3}$  & $2$    	\\
      \\
    \end{tabular}
%\end{adjustwidth}
  %\end{center}
\end{minipage}
\end{table}

@@

Clustering algorithms are unsupervised learning methods whose goal is to partition a set of objects into (disjoint) subsets, called clusters, such that each cluster contains the most similar objects. 
The similarity of objects is determined by some measure, like Euclidian distance, using features of each object, such as features of operations extracted in the previous section. 
However, direct application of clustering algorithms, like K-Means \cite{Forgy1965ClusterAO}, in scheduling is limited since the partitioning cannot preserve the sequences of operations in a job. As a result, a clustering algorithm might place $O_{1,1}$ and $O_{1,3}$ in one cluster and $O_{1,2}$ in another. As a result, the sequence of sub-problems becomes inconsistent and no schedule can be found.

@@

\section{Related Work}
\label{sec:literature}
In this section, we review some of the most related work using data mining approaches employed to solve scheduling problems. Over the last decades, solving the scheduling problem has attracted the interest of many researchers. The main idea of data mining in these applications is to explore the patterns from the obtained solutions by some heuristics and then develop decision rules that approximate the order of the operations in each machine \cite{ismail2012production}. More precisely, it aims to get an order of all operations in the given instance according to their characteristics.
